---
title: "Figure 1"
author: David McGaughey
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: 
  html_notebook:
    theme: flatly
    toc: true
    code_folding: hide
---
```{r}
library(tidyverse)
library(ggforce)
library(ggrepel)
library(cowplot)
load('/Volumes/data/projects/nei/mcgaughey/massive_integrated_eye_scRNA/umap_integrated_obj_scran_fastMNN_200_res15.Rdata')
load('/Volumes/data/projects/nei/mcgaughey/massive_integrated_eye_scRNA/umap/Mus_musculus_Macaca_fascicularis_Homo_sapiens__standard__full__batch__fastMNN__dims50.umap.Rdata')

# attach colors to cell types
cell_types <- umap %>% mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>% 
  pull(CellType) %>% unique() %>% sort()
type_val <- setNames(pals::alphabet(n = cell_types %>% length()), cell_types)
type_col <- scale_colour_manual(values = type_val)
type_fill <- scale_fill_manual(values = type_val)
# 
cluster_labels <- umap %>%
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>%
  group_by(seurat_clusters, CellType) %>%
  filter(!is.na(CellType), !is.na(seurat_clusters)) %>%
  summarise(Count = n(), x = mean(UMAP_1), y = mean(UMAP_2), 
            Organism = list(unique(organism)), 
            study_accession = list(unique(study_accession))) %>%
  mutate(freq = Count / sum(Count)) %>%
  filter(freq > 0.25) %>%
  ungroup() %>%
  group_by(seurat_clusters) %>%
  top_n(3, -freq) %>%
  #filter(Count > 100) %>%
  summarise(seurat_cluster_CellType = paste0(CellType, collapse = ', '),
            x = mean(x), y = mean(y),
            Count = sum(Count),
            OrganismCount = unlist(Organism) %>% unique() %>% sort() %>% paste(., collapse = ', '),
            StudyCount = unlist(study_accession) %>% length()) %>%
  mutate(seurat_cluster_CellType_num = paste0(seurat_clusters, ' (', seurat_cluster_CellType, ')'))
```

# UMAP of known cells
```{r, fig.width=6, fig.height=4}
plot1 <- umap %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  #left_join(., cluster_labels) %>% 
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.05, alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num ), alpha = 0.8, size = 2) +
  type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  #coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
plot1  #+ facet_wrap(~CellType)
# pdf('big_umap_fig2.pdf', width = 12, height = 9)
# plot1
# dev.off()
```

# UMAP of Age
```{r, fig.width=4, fig.height=4}
plot2 <- umap %>% 
  rename(Stage = integration_group) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = Stage), size = 0.01, alpha = 0.01) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  scale_color_manual(values = as.vector(pals::alphabet()) %>% tail(2)) + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
plot2
```

# UMAP of inferred cells, split by organism
```{r, fig.width=9, fig.height=4}
plot3 <- umap %>% 
  mutate(CellType_predict = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(study_accession), !is.na(CellType_predict),
         !CellType_predict %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType_predict)) %>%
  mutate(Size = case_when(organism == 'Homo sapiens' ~ 0.015,
                          TRUE ~ 0.01)) %>% 
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType_predict, size = Size), alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=10, alpha = 1))) + 
  theme_cowplot() + 
  scale_size(guide = 'none') +
  scale_alpha(guide = 'none') +
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) + 
  facet_wrap(~organism) +
  xlab('UMAP 1') + ylab('UMAP 2')
plot3
```

```{r}
pdf('big_umap.pdf', height = 10, width = 10)
plot_grid(plot1, plot_grid(plot3 + theme(legend.position = "none"), plot2, 
                           rel_widths = c(0.65,0.35)), 
          nrow = 2, 
          rel_heights = c(0.6,0.4))
dev.off()
```

# Amacrine
```{r, fig.width=6, fig.height=4}
umap %>% 
  filter(GAD1 > 0) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.05, alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
```

# Bipolar
```{r, fig.width=6, fig.height=4}
umap %>% 
  filter(OPN1SW > 0) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.05, alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  scale_color_manual(values = as.vector(pals::alphabet())) + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
```

# Bipolar Cell Analysis
```{r}
# bc <- FindMarkers(integrated_obj, ident.1 = c(10,43,33), ident.2 = c(27,97))
plotX <- umap %>% 
  left_join(., cluster_labels, by = 'seurat_clusters') %>% 
  filter(seurat_cluster_CellType == 'Bipolar Cells') %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot()  +
  #type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  #coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')

plot_grid(plotX + geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.5, alpha = 0.1) + type_col + theme(legend.position = "none"), 
          plotX + geom_point(aes(x=UMAP_1, y = UMAP_2, colour = seurat_clusters), size = 0.5, alpha = 0.1) +
            geom_label_repel(data = cluster_labels %>% 
                               filter(seurat_cluster_CellType == 'Bipolar Cells'), 
                             aes(x=x, y=y, label = seurat_clusters), alpha = 0.8, size = 4) +
            theme(legend.position = "none") , 
          align = 'v',
          ncol = 1)

# violin plot of rod bc vs cone bc markers
# cabp5 is a general bc marker
umap %>% 
    left_join(., cluster_labels, by = 'seurat_clusters') %>% 
    filter(seurat_cluster_CellType == 'Bipolar Cells') %>% 
    mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
    filter(!is.na(CellType), 
           !is.na(study_accession), 
           !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
           !grepl('RPE|Vascul', CellType)) %>% 
    mutate(Group = case_when(seurat_clusters %in% c(20, 34) ~ 'BC1 (20, 34)',
                             seurat_clusters %in% c(41,51) ~ 'BC2 (41, 51)')) %>% gather(Gene, Expression, RHO:SCGN) %>% 
  filter(!is.na(Group), Gene %in% c('CABP5', 'VSX2', 'PRKCA')) %>% 
  ggplot() + geom_violin(aes(x=Group, y = Expression, colour = Gene)) + theme_cowplot()




# SCNG expression
# scng is amarine/RGC marker?????
# plot is filtered down only to labelled Bipolar Cells
umap %>% 
    left_join(., cluster_labels, by = 'seurat_clusters') %>% 
   mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
    filter(seurat_cluster_CellType == 'Bipolar Cells') %>% 
   
    filter(!is.na(CellType), CellType == 'Bipolar Cells',
           !is.na(study_accession), 
           !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
           !grepl('RPE|Vascul', CellType)) %>%
    ggplot() + 
    guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
    theme_cowplot()  +
    #type_col + 
    theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
    #coord_cartesian(xlim = c(-5,5)) +
    xlab('UMAP 1') + ylab('UMAP 2') + geom_point(aes(x=UMAP_1, y = UMAP_2, colour = SNCG), size = 0.5, alpha = 0.1) + scale_color_viridis_c() 
```


# RGC Analysis
```{r}
# bc <- FindMarkers(integrated_obj, ident.1 = c(10,43,33), ident.2 = c(27,97))
plotX <- umap %>% 
  left_join(., cluster_labels, by = 'seurat_clusters') %>% 
  filter(seurat_cluster_CellType == 'Retinal Ganglion Cells') %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot()  +
  #type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  #coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')

plot_grid(plotX + geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.5, alpha = 0.1) + type_col + theme(legend.position = "none"), 
          plotX + geom_point(aes(x=UMAP_1, y = UMAP_2, colour = seurat_clusters), size = 0.5, alpha = 0.1) +
            geom_label_repel(data = cluster_labels %>% 
                               filter(seurat_cluster_CellType == 'Retinal Ganglion Cells'), 
                             aes(x=x, y=y, label = seurat_clusters), alpha = 0.8, size = 4) +
            theme(legend.position = "none") , 
          align = 'v',
          ncol = 1)
```

Clustering Cell Type Analysis
` r umap %>% filter(!is.na(CellType), !CellType %in% c('Neurogenic Cells', 'Photoreceptor Precursors', 'Late RPCs', 'Early RPCs'), integration_group == 'Late') %>% nrow()` total cells excluding the Clark et al. specific cell types and early development cells.


```{r}
umap %>%
   filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>%
  group_by(seurat_clusters, CellType) %>%
  filter(!is.na(CellType), !is.na(seurat_clusters)) %>%
  summarise(Count = n(), x = mean(UMAP_1), y = mean(UMAP_2), 
            Organism = list(unique(organism)), 
            StudyCount = length(unique(study_accession))) %>%
  mutate(freq = Count / sum(Count)) %>%
  filter(freq > 0.25, Count > 100) %>% 
  ggplot(aes(x=CellType, y = StudyCount)) + 
  coord_flip() + 
  geom_boxplot(outlier.colour = NA) + 
  ggbeeswarm::geom_quasirandom(alpha = 0.3) +
  ylab('Study Count') + xlab('Cell Type') +
  theme_cowplot()
```