---
title: "Figure 1"
author: David McGaughey
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: 
  html_notebook:
    theme: flatly
    toc: true
    code_folding: hide
---
```{r}
library(tidyverse)
library(ggforce)
library(ggrepel)
library(cowplot)
load('/Volumes/data/projects/nei/mcgaughey/massive_integrated_eye_scRNA/umap_fastMNN__mindist001__pc200_2.Rdata')
# attach colors to cell types
cell_types <- umap %>% mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>% 
  pull(CellType) %>% unique() %>% sort()
type_val <- setNames(pals::alphabet(n = cell_types %>% length()), cell_types)
type_col <- scale_colour_manual(values = type_val)
type_fill <- scale_fill_manual(values = type_val)
# 
# cluster_labels <- umap %>% 
#   mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
#   group_by(seurat_clusters, CellType) %>% 
#   filter(!is.na(CellType)) %>% 
#   summarise(Count = n(), x = mean(UMAP_1), y = mean(UMAP_2)) %>% 
#   mutate(freq = Count / sum(Count)) %>% 
#   filter(freq > 0.2) %>% 
#   ungroup() %>% 
#   group_by(seurat_clusters) %>% 
#   top_n(1, -freq) %>% 
#   filter(Count > 100) %>% 
#   summarise(seurat_cluster_CellType = paste0(CellType, collapse = ', '),
#             x = mean(x), y = mean(y)) %>% 
#   mutate(seurat_cluster_CellType_num = paste0(seurat_clusters, ' (', seurat_cluster_CellType, ')'))
```

# UMAP of known cells
```{r, fig.width=6, fig.height=4}
plot1 <- umap %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.05, alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
plot1  #+ facet_wrap(~CellType)
# pdf('big_umap_fig2.pdf', width = 12, height = 9)
# plot1
# dev.off()
```

# UMAP of Age
```{r, fig.width=4, fig.height=4}
plot2 <- umap %>% 
  rename(Stage = integration_group) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = Stage), size = 0.01, alpha = 0.01) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  scale_color_manual(values = as.vector(pals::alphabet()) %>% tail(2)) + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
plot2
```

# UMAP of inferred cells, split by organism
```{r, fig.width=9, fig.height=4}
plot3 <- umap %>% 
  mutate(CellType_predict = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(study_accession), !is.na(CellType_predict),
         !CellType_predict %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType_predict)) %>%
  mutate(Size = case_when(organism == 'Homo sapiens' ~ 0.015,
                          TRUE ~ 0.01)) %>% 
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType_predict, size = Size), alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=10, alpha = 1))) + 
  theme_cowplot() + 
  scale_size(guide = 'none') +
  scale_alpha(guide = 'none') +
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) + 
  facet_wrap(~organism) +
  xlab('UMAP 1') + ylab('UMAP 2')
plot3
```

```{r}
pdf('big_umap.pdf', height = 10, width = 10)
plot_grid(plot1, plot_grid(plot3 + theme(legend.position = "none"), plot2, 
                           rel_widths = c(0.65,0.35)), 
          nrow = 2, 
          rel_heights = c(0.6,0.4))
dev.off()
```

# Amacrine
```{r, fig.width=6, fig.height=4}
umap %>% 
  filter(GAD1 > 0) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.05, alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  type_col + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
```

# Bipolar
```{r, fig.width=6, fig.height=4}
umap %>% 
  filter(OPN1SW > 0) %>% 
  mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets', 'Fibroblasts', 'Red Blood Cells'),
         !grepl('RPE|Vascul', CellType)) %>%
  ggplot() + 
  geom_point(aes(x=UMAP_1, y = UMAP_2, colour = CellType), size = 0.05, alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(size=8, alpha = 1))) + 
  theme_cowplot() + 
  #geom_label_repel(data = cluster_labels, aes(x=x, y=y, label = seurat_cluster_CellType_num )) +
  scale_color_manual(values = as.vector(pals::alphabet())) + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  coord_cartesian(xlim = c(-5,5)) +
  xlab('UMAP 1') + ylab('UMAP 2')
```

umap %>% left_join(., cluster_labels) %>% 
  mutate(Time = integration_group,
         CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType), 
         Labelling = case_when(Paper == 'Hufnagel 2020' ~ 'Homo sapiens (Stem) (Hufnagel 2020)',
                               TRUE ~ paste0(organism, ' (', study_accession, ')'))) %>% 
  sample_n(10000) %>% 
  ggplot(aes(x=mnnUMAP_1, y = mnnUMAP_2)) + 
  geom_point(size = 0.1, alpha = 0.5) + 
  guides(colour = guide_legend(override.aes = list(size=10, alpha = 1))) + 
  theme_cowplot() + 
  geom_mark_hull(aes(label = seurat_cluster_CellType_num), concavity = 5) +
  scale_color_manual(values = as.vector(pals::alphabet())) + 
  scale_fill_manual(values = as.vector(pals::alphabet())) + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
```

