---
title: "Optimal Params"
output: html_notebook
---

Pick optimal knn, nfeatures, nneighbors, dims based upon:

  - ARI (adjusted rand index, ratio of agreements over disagreements. 1 perfectly agrees, 0 perfectly disagrees. Higher is better, as the ARI is calculted as agreement between cluster and known cell type labels)
  - Silhouette (scVI dimensions against [Batch, CellType, Cluster]). A score of -1 means perfect mixing, 1 means perfectly non-mixed (good for Cluster, CellType)
  - LISI (scVI dimensions against [Batch, CellType, Cluster]). A score of 1 means that each cell has, on average, one [Batch, CellType, Cluster] near it. Higher is better for Batch. Closest to 1 is best for CellType, Cluster.
    - Silhouette and LISI each have three scores: 
      - Batch 
      - CellType 
      - Cluster 
    
```{r}
library(tidyverse)
library(cowplot)
load('/Volumes/OGVFB_BG/scEiaD/metrics_2020_08_20.Rdata')


perf_well <- perf %>% unique() %>% filter(set == 'onlyWELL') %>% 
  select(Score, Group, Value, set, dims:normalization) %>% 
  filter(Score %in% c('LISI','Silhouette')) 


perf_uni <- perf %>% unique() %>% filter(set == 'universe') %>% 
  select(Score, Group, Value, dims:normalization) %>% 
  filter(Score %in% c('LISI','Silhouette', 'ARI', 'PCR', 'NMI')) 


perf <- perf %>% unique() %>% filter(set == 'TabulaDroplet') %>% 
  select(Score, Group, Value, set, dims:normalization) %>% 
  filter(Score %in% c('LISI','Silhouette', 'ARI', 'PCR', 'NMI')) 


perf %>% DT::datatable()
```


# LISI Metric
Want Batch as high as possible (mean # of batches near each cell) and Cluster as low as possible (mean # of cluster types near each cell)
```{r}
perf %>% 
  filter(Score == 'LISI', nf == 2000, dims %in% c(10,30), knn == 7) %>%
  pivot_wider(values_from = Value, names_from = c('Group')) %>% 
  ggplot(aes(y=Batch, x=-Cluster, shape = normalization)) + geom_point(aes(color=method), size = 5) +
  cowplot::theme_cowplot() + scale_color_manual(values = pals::alphabet() %>% unname())
```

# ARI
```{r}
perf %>% 
  filter(Score == 'ARI', Group == 'CellType-Cluster', nf == 2000, dims %in% c(10, 30), knn == 7) %>%
  ggplot(aes(y=Value, x=normalization)) + geom_point(aes(color=method), size = 5) +
  cowplot::theme_cowplot() + scale_color_manual(values = pals::alphabet() %>% unname()) + ylab('ARI (higher is better)')
```

# Silhouette Metric
Want Batch as low as possible (-1 is perfect mixing) and Cluster as high as possible (1 is perfect purity of cluster types near each cell). Notice how the y-axis has been "flipped" by multiply by -1.
```{r}
perf %>% filter(Score == 'Silhouette', nf == 2000, dims %in% c(10,30), knn == 7) %>% 
    pivot_wider(values_from = Value, names_from = c('Group')) %>% 
  ggplot(aes(y=-Batch, x=Cluster, shape = normalization)) + geom_point(aes(color=method), size = 5) +
  cowplot::theme_cowplot() + scale_color_manual(values = pals::alphabet() %>% unname())
```

# Table
## TabulaDroplet
Use Z scaling to get `sumZScale` score where the highest value is, overall across all the metrics, the best performer
```{r}
perf %>% 
  pivot_wider(names_from = c('Score','Group'), values_from = Value) %>% 
  mutate(sumZScale = 
           -scale(LISI_CellType)[,1] +
           -scale(LISI_SubCellType)[,1] +
           scale(Silhouette_CellType)[,1] +
           #scale(Silhouette_SubCellType)[,1] +
           scale(`NMI_CellType-Cluster`)[,1] +
           #scale(`NMI_SubCellType-Cluster`)[,1] +
           scale(`ARI_CellType-Cluster`)[,1] +
           #scale(`ARI_SubCellType-Cluster`)[,1] +
           
           scale(LISI_Batch)[,1] + # Z score
           -scale(LISI_Cluster)[,1] +
           -scale(Silhouette_Batch)[,1] + # Z score
           scale(Silhouette_Cluster)[,1] +
           scale(`PCR_After-Before`)[,1]
           ) %>% 
  arrange(-sumZScale) %>% DT::datatable()
```

## onlyWELL
Use Z scaling to get `sumZScale` score where the highest value is, overall across all the metrics, the best performer
```{r}
perf_well %>% 
  pivot_wider(names_from = c('Score','Group'), values_from = Value) %>% 
  mutate(sumZScale = 
         
           scale(LISI_Batch)[,1] + # Z score
           2 * -scale(LISI_Cluster)[,1] + 
           -scale(Silhouette_Batch)[,1] + # Z score
           2 * scale(Silhouette_Cluster)[,1] 
           ) %>% 
  arrange(-sumZScale) %>% DT::datatable()
```
## Universe 
(well and droplet and tabula muris)
```{r}
perf_uni %>% 
  pivot_wider(names_from = c('Score','Group'), values_from = Value) %>% 
  mutate(sumZScale = 
           -scale(LISI_CellType)[,1] +
           #-scale(LISI_SubCellType)[,1] +
           scale(Silhouette_CellType)[,1] +
           #scale(Silhouette_SubCellType)[,1] +
           scale(`NMI_CellType-Cluster`)[,1] +
           #scale(`NMI_SubCellType-Cluster`)[,1] +
           scale(`ARI_CellType-Cluster`)[,1] +
           #scale(`ARI_SubCellType-Cluster`)[,1] +
           
           scale(LISI_Batch)[,1] + # Z score
           -scale(LISI_Cluster)[,1] +
           -scale(Silhouette_Batch)[,1] + # Z score
           scale(Silhouette_Cluster)[,1] +
           scale(`PCR_After-Before`)[,1]
           ) %>% 
  arrange(-sumZScale) %>% DT::datatable()
```

```{r}
devtools::session_info()
#save.image(file = paste0('../data/',Sys.Date(), 'optimal_params.Rdata'))
```






