---
title: "Download New Samples"
output: html_notebook
---

Date searched: 12/07/2020
- Search GEO DataSet for "(((single[Description]) AND cell[Description]) AND ((retina[Description]) OR RPE[Description])) AND (((Homo sapiens[Organism]) OR Mus Musculus[Organism]) OR Macaca[Organism])"
  - search GEO and not SRA because theres no publication data filter on SRA
"
- set publication date to within 1 yr
- Find related data > SRA 
- send to file  > full xml 
- send to runSelector > download metadata


```{r}
library(tidyverse)
```

```{r}
library(xml2)
library(tidyverse)
full_xml <- read_xml('~/Downloads/SraExperimentPackage-4.xml') %>% as_list()
curr_sampletable <- read_tsv('../data/sample_run_layout_organism_tech.tsv')
curr_studies <- curr_sampletable$study_accession %>% unique
exp_set <- full_xml$EXPERIMENT_PACKAGE_SET
exp_set_list <- lapply(exp_set, function(x) unlist(x, recursive = T))
common_names <- lapply(exp_set_list, names) %>% reduce(intersect)
res <-lapply(exp_set_list , function(x) x[common_names]) %>%   do.call(rbind,.) %>% as.data.frame()
colnames(res)
study_cols <- c("EXPERIMENT.STUDY_REF.IDENTIFIERS.PRIMARY_ID", "STUDY.DESCRIPTOR.STUDY_TITLE" ,"STUDY.DESCRIPTOR.STUDY_ABSTRACT"  )
res_by_study <- res %>% select(all_of(study_cols)) %>% distinct
metadata <- read_csv('~/Downloads/SraRunTable.txt-4.csv')
res_by_study_with_date <- metadata %>% select( EXPERIMENT.STUDY_REF.IDENTIFIERS.PRIMARY_ID=`SRA Study`, ReleaseDate) %>% 
  distinct %>% left_join(res_by_study, .) %>% 
  rename(study_accession =EXPERIMENT.STUDY_REF.IDENTIFIERS.PRIMARY_ID)
new_studies <- filter(res_by_study_with_date, !study_accession %in% curr_studies)

write_csv(new_studies, 'new_potential_Studies.csv')
```
 
what is Tamoxifen?
 - induces the CRE transgene

SRP238409 - only Donor 4 and 5 deposited? look like the damaged ones


```{r}
keep_samples <- read_csv('new_potential_Studies_selected.csv') 

# use grepl to ignore NAs
metadata_keep_samples <-  filter(metadata, `SRA Study` %in% keep_samples$study_accession, 
       !grepl('mut|Mutant|KO|knockout|treated|-/-|Lung|Liver|Heart|GLI1\\+', source_name), 
       !grepl('Light damage|NMDA treatment', Condition), 
       !grepl('Danio rerio', Organism), 
       !grepl('Ihhi', `genotype/variation`),
       !grepl('Donor 5', Donor),
       !grepl('L75Pfs|/\\+|Nrl-L-EGFP', Genotype),
       !grepl('Retinal ganglion cell deficiency', PHENOTYPE)
       ) %>% 
  rename(study_accession =  `SRA Study`, 
         )
kcols <- apply(metadata_keep_samples, 2, function(x) all(is.na(x)))
metadata_keep_samples <- metadata_keep_samples[,!kcols]
cols <- c('preparation', 'Donor', 'Individual', 'Replicate')
cov_df <-  lapply(cols, function(x) select(metadata_keep_samples, Run, Covariate = !!x ) %>% 
                    filter(!is.na(Covariate))) %>% bind_rows
sum(duplicated(cov_df$Run))
new_sample_metadata <-  left_join(metadata_keep_samples, cov_df) %>% 
  select(sample_accession = Experiment, 
         run_accession = Run, 
         library_layout = LibraryLayout,
         organism=Organism, 
         study_accession,
         msg_tissue=Tissue, 
         Covariate,
         Age=AGE,
         Cell_type,
         source_name,
         position
         ) %>% arrange(study_accession) %>% 
  mutate(Tissue = case_when(
    grepl('retina', msg_tissue, ignore.case = T) ~ 'Retina',
    grepl('retina', Cell_type)~ 'Retina',
    grepl('RPE', Cell_type) ~ 'RPE',
    grepl('retina',source_name, ignore.case = T) ~ 'Retina'
    
  )) %>% unite('TissueNote', Cell_type:position, sep = ';') %>% 
  mutate(TissueNote = str_remove_all(TissueNote,'^NA;|;NA;|;NA$'), 
         UMI='YES', 
         integration_group = '12/09/2020', 
         Covariate = replace(Covariate, is.na(Covariate), study_accession[is.na(Covariate) ])) %>% 
  inner_join(keep_samples %>% select(study_accession,Platform)) %>% 
  select(all_of(colnames(curr_sampletable)))

colnames(metadata_keep_samples)
write_tsv(new_sample_metadata, '../data/20201209_new_sample_metadata.tsv')

most_curr_sample_table <- read_tsv('https://raw.githubusercontent.com/davemcg/scEiaD/velocity/data/sample_run_layout_organism_tech_20201030.tsv')

full_sample_table <- bind_rows(most_curr_sample_table,new_sample_metadata %>% mutate(Age=1000000))
write_tsv(full_sample_table, '../data/sample_run_layout_organism_tech_20201210.tsv')
```




