---
title: "Table 1"
author: David McGaughey
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: 
  html_notebook:
    theme: flatly
    toc: true
    code_folding: hide
---

```{r, message=F, warning=F, include=F}
# Load Libraries without printing any warnings or messages
library(tidyverse)
library(formattable)
library(webshot)
library(htmltools)
library(pool)
library(RSQLite)
# all cells
load('/Volumes/data/projects/nei/mcgaughey/massive_integrated_eye_scRNA/cell_info_labelled.Rdata')

# load('/Volumes/data/projects/nei/mcgaughey/massive_integrated_eye_scRNA/fastMNN_umap_full.Rdata')
study_meta <- read_tsv('~/git/massive_integrated_eye_scRNA/data/GEO_Study_Level_Metadata.tsv')

scEiaD <- dbPool(drv = SQLite(), dbname = "~/data/massive_integrated_eye_scRNA/MOARTABLES__anthology_limmaFALSE___Mus_musculus_Macaca_fascicularis_Homo_sapiens-5000-counts-TabulaDroplet-batch-scVI-8-0.1-15-7.sqlite", idleTimeout = 3600000)
meta <- scEiaD %>% tbl('metadata') %>% as_tibble()

stats <- meta %>% group_by(study_accession, Platform) %>% summarise(Counts = n())
```

# Table 1
Information on number/composition of studies

`r nrow(cell_info_labels)` cells enter the ring.

`r cell_info_labels$study_accession %>% unique() %>% length()` SRA studies

`r table01$Citation %>% unique() %>% length() + table01 %>% filter(Citation == '') %>% nrow()` unique studies

```{r}
color_bar_factor <- formatter("span",
                              style = function(x) style(
                                display = "block",
                                color = "white",
                                border.radius = "4px",
                                background = c("#E69F00", "#56B4E9", "#009E73")[factor(as.character(x))]))

color_bar_factor2 <- formatter("span",
                              style = function(x) style(
                                display = "block",
                                color = "white",
                                border.radius = "4px",
                                background = c("#0072B2", "#D55E00")[factor(as.character(x))]))

color_bar_factor3 <- formatter("span", width = "5px",
                              style = function(x) style(
                                display = "block",
                                color = "white",
                                border.radius = "4px",
                                background = type_val[factor(as.character(x))]))

export_formattable <- function(f, file, width = "100%", height = NULL, 
                               background = "white", delay = 0.2)
{
  w <- as.htmlwidget(f, width = width, height = height)
  path <- html_print(w, background = background, viewer = NULL)
  url <- paste0("file:///", gsub("\\\\", "/", normalizePath(path)))
  webshot(url,zoom = 4,
          file = file, 
          selector = ".formattable_widget",
          delay = delay)
}

post <- stats %>% rename(`Post QC<br/>Count` = Counts) %>% 
              select(study_accession, Platform, `Post QC<br/>Count`) %>% 
              group_by(study_accession, Platform) %>% 
              summarise(`Post QC<br/>Count` = sum(`Post QC<br/>Count`))

table01 <- cell_info_labels %>% 
  filter(!is.na(study_accession)) %>% 
  left_join(study_meta) %>% 
  mutate(PMID = as.character(PMID)) %>% 
  mutate(Citation = case_when(is.na(Citation) ~ '',
                              TRUE ~ paste0(substr(Citation, 1, 30), ' ...')),
         PMID = case_when(is.na(PMID) ~ '',
                          TRUE ~ PMID),
         `SRA Accession` = study_accession,
         Labels = case_when(!is.na(CellType) ~ 'Yes',
                              TRUE ~ 'No')) %>% 
  group_by(Citation, PMID, `SRA Accession`, organism, Platform) %>% 
  summarise(Count = n(), Labels = unique(Labels) %>% sort() %>% tail(1)) %>% 
  arrange(organism, -Count) %>% 
  left_join(., post %>% 
              rename(`SRA Accession` = study_accession)) %>% 
  select(Citation:Count, `Post QC<br/>Count`, Labels)

table01$`Post QC<br/>Count`[is.na(table01$`Post QC<br/>Count`)] <- 0
formattable_01 <- table01 %>% 
  formattable(., list(Count = normalize_bar("lightblue"),
                      `Post QC<br/>Count` = normalize_bar("lightblue"),
                      organism = color_bar_factor,
                      Labels = color_bar_factor2,
                      Platform = formatter("span", style = x ~ ifelse(!grepl('SMART|C1', x), 
                                                                      style(font.weight = "bold"), NA))))
formattable_01
# export_formattable(formattable_01, '01_table.png')
write_file(format_table(formattable_01), path = 'table_01.html')
```
`r table01 %>% filter(Labels == 'Yes') %>% nrow()` studies pre(labelled).

`r sum(table01$"Post QC<br/>Count")` cells after QC.

# Table 2
Pre-labelled cells I've extracted so far
```{r}
table02 <- cell_info_labels %>% 
  mutate(CellType = gsub('AC/HC_Precurs', 'Amacrine/Horizontal Precursors', CellType)) %>% 
  #mutate(CellType = gsub("Rod Bipolar Cells", "Bipolar Cells", CellType)) %>% 
  filter(!is.na(CellType), 
         !is.na(study_accession), 
         !CellType %in% c('Doublet', 'Doublets'),
         !grepl('RPE|Vascul', CellType)) %>% 
  mutate(organism = case_when(grepl('Homo', organism) ~ 'HS',
                             grepl('Mus', organism) ~ 'MM',
                             TRUE ~ 'MF')) %>% 
  group_by(CellType,organism, study_accession) %>% 
  summarise(Count = n()) %>% 
  summarise(Studies = length(study_accession), Count = sum(Count)) %>% 
  summarise(Species = paste(organism, collapse = ', '), Studies = sum(Studies), Count = sum(Count)) %>% 
  arrange(-Count) %>% 
  formattable(., list(Count = normalize_bar("lightblue")), width = "1")
table02
formattable_02 <- table02
write_file(format_table(formattable_02), path = 'table_02.html')

save(formattable_01, formattable_02, file = 'formattables.Rdata')
```

# Table 3 
Alluvial plot of Labelled CellTypes and Projected CellTypes
```{r}
library(ggalluvial)

meta %>% 
  mutate(CellType = case_when(is.na(CellType) ~ 'Unlabelled', TRUE ~ CellType)) %>% 
  select(CellType, CellType_predict) %>% 
  group_by(CellType, CellType_predict) %>% 
  summarise(Count = n()) %>% 
  ggplot(aes(y = log2(Count), axis1 = CellType, axis2 = CellType_predict, )) +
  geom_alluvium(aes(fill = `CellType`), alpha = 0.8) +
  geom_stratum(alpha = 0) +
  # ggrepel::geom_text_repel(stat = "stratum", 
  #                          label.strata = TRUE, 
  #                          direction = "y",
  #                          nudge_x = c(rep(c(-10),32), rep(c(10),32))) +
  ggfittext::geom_fit_text(stat = "stratum", width = 1/4, min.size = 2,
                           aes(label = CellType_predict)) +
  ggfittext::geom_fit_text(stat = "stratum", width = 1/4, min.size = 2,
                           aes(label = CellType)) +
  coord_cartesian(xlim = c(0.5,2.5)) +
  theme_void() +
  scale_fill_manual(values = c(pals::polychrome() %>% unname(),
                    pals::alphabet() %>% unname())) +
   theme(legend.title = element_blank())


```
# Table 4
Number of annotated cells in well supported clusters
```{r}
# load('/Volumes/data/projects/nei/mcgaughey/massive_integrated_eye_scRNA/umap/Mus_musculus_Macaca_fascicularis_Homo_sapiens__n_features5000__counts__full__batch__scVI__dims30__mindist0.3__nneighbors50.umap.Rdata')
meta %>%
  #mutate(CellType = gsub('Rod Bipolar Cells', 'Bipolar Cells', CellType)) %>%
  mutate(CellType = CellType_predict) %>%
  group_by(cluster, CellType) %>%
  filter(!is.na(CellType), !is.na(cluster)) %>%
  summarise(Count = n(), x = mean(UMAP_1), y = mean(UMAP_2),
            Organism = list(unique(organism)),
            study_accession = list(unique(study_accession))) %>%
  mutate(freq = Count / sum(Count)) %>%
  filter(freq > 0.25) %>%
  ungroup() %>%
  group_by(cluster) %>%
  top_n(3, -freq) %>%
  #filter(Count > 100) %>%
  summarise(seurat_cluster_CellType = paste0(CellType, collapse = ', '),
            x = mean(x), y = mean(y),
            Count = sum(Count),
            OrganismCount = unlist(Organism) %>% unique() %>% sort() %>% paste(., collapse = ', '),
            StudyCount = unlist(study_accession) %>% length()) %>%
  mutate(seurat_cluster_CellType_num = paste0(cluster, ' (', seurat_cluster_CellType, ')')) %>%
  select(seurat_cluster_CellType_num, OrganismCount, StudyCount, Count) %>%
  mutate(OrganismCount = gsub('Homo sapiens', 'HS', OrganismCount)) %>%
  mutate(OrganismCount = gsub('Macaca fascicularis', 'MF', OrganismCount)) %>%
  mutate(OrganismCount = gsub('Mus musculus', 'MM', OrganismCount)) %>%
  dplyr::rename(Cluster = seurat_cluster_CellType_num, Species = OrganismCount, `Study Count`  = StudyCount) %>%
  #filter(`StudyCount` > 4) %>%
  arrange(-Count) %>%
  formattable(., list(Count = normalize_bar("lightblue")), width = "0.5")

```
# Session Info
```{r}
devtools::session_info()
```